Loading data frame of PFizer Company data. We will be carrying out the analysis.

```{r}
library(forecast)
library(tseries)
library(fpp2)
library(ggplot2)
library(dplyr)
library(lubridate)
library(plotly)
library(quantmod)
library(tidyquant)
```


```{r}
#Taking the path of the csv file.

Path = "/Applications/Fall Semester/Homework/Linear regression/Project/PFE.csv"

#Loading the dataframe
PFizer <-read.csv("PFE.csv" ,header = TRUE,na.strings = 999.99)
```

Printing out the head of the PFizer data frame
```{r}
head(PFizer)
```
Printing out the tail of the PFizer data frame
```{r}
tail(PFizer)
```
Loading the data frame in tabular format
```{r}
View(PFizer)
```
Loading the dimension of the data frame
```{r}
dim(PFizer)
```
The data frame has 946 rows and 7 columns

View structure of data
```{r}
str(PFizer)
```

Converting the date column data type from chr to date

```{r}
PFizer <- PFizer %>% mutate(Date = as.Date(Date, format = "%d/%m/%Y"))
```

Adding column month to the data frame
```{r}
PFizer <- PFizer %>% mutate(month=month(Date))
```

Adding column year to the data frame
```{r}
PFizer <- PFizer %>% mutate(year=year(Date))
```

Loading the new data frame
```{r}
head(PFizer)
```
We will study the trend of mean_close parameter of the data frame.

Lets create a line plot of the full time series of PFizer sector stocks, using Dataframe's plot() method
```{r}
plot(PFizer$year,PFizer$Close,xlab="Month",ylab="Close",type="l",lwd=2,col="blue")
```
To get a better view of trend, let's plot the data frame w.r.t mean_close for the time period.
```{r}
# plot the data using ggplot2 and pipes
PFizer %>%
ggplot(aes(x = Date, y = Close)) +
      geom_point(color = "darkorchid4") +
      labs(title = "Closing trend - PFizer",
           subtitle = "The data frame is sent to the plot using pipes",
           y = "Closing trend",
           x = "Date") + theme_bw(base_size = 15)
```
We can get more insight from the data by plotting it using logs
Taking log values of closing values and the differences of logs
```{r}
plot(10*diff(log(PFizer[,6])),xlab="year",ylab="Close",type="l",lwd=2,col="orange")
```
Let's see for year 2022
```{r}
year_data_PFizer=subset(PFizer,subset=PFizer$Date>='2022-01-01' & PFizer$Date<='2022-12-31')
head(year_data_PFizer)
```
```{r}
plot(year_data_PFizer[,1],year_data_PFizer[,6],xlab="Year 2022",ylab="Close",type="l",lwd=2,main="Close values",col="orange")
```
We see, that the trend is starting with high and then decreasing and increasing continuously for the year 2022.

Let's do box plot for yearly data.
Taking quantile of the mean_close
```{r}
quantile(PFizer$Close,probs=c(0,0.25,0.5,0.75,1))
```
plotting the boxplot
```{r}
boxplot(PFizer$Close~PFizer$year,main="Closing trend over the years 2019-2022",ylab="Closing",xlab="years",las=1,col="green")
```
We see that the year 2022 is performing better than rest of the years..

Now, let's do the analysis based on months for the years 2019-2022

```{r}
# plot the data using ggplot2 and pipes
PFizer %>%
ggplot(aes(x = month, y = Close)) +
      geom_point(color = "darkorchid4") +
      labs(title = "Monthly Closing trend - PFizer",
           subtitle = "The data frame is sent to the plot using pipes",
           y = "Closing",
           x = "Date") + theme_bw(base_size = 15)

```
Here, we see that the monthly trend is not that clear as the month trend for one year is getting overlapped with another year.

Let's try another graph which is line graph to see if we can see something clearly compared to previous graph
```{r}
ggplot(data=PFizer,aes(x=month,y=Close,col="red",group=1))+geom_line(linetype="dashed")+geom_point()
```
Again, we cannot infer clearly here.

To overcome this problem, let's group the data by month and year
```{r}
# calculate the sum closing for each month
PFizer_month <- PFizer %>%
  group_by(month, year) %>%
  summarise(max_Close = sum(Close))
```

Loading the newly created dataset where the data is 
```{r}
head(PFizer_month)
```
plotting the new data frame to see the trend

```{r}
PFizer_month %>%
  mutate(month2 = as.Date(paste0("2019-", month,"-01"),"%Y-%m-%d")) %>%
  ggplot(aes(x = month2, y = max_Close)) +
      geom_bar(stat = "identity", fill = "darkorchid4") +
  facet_wrap(~ year, ncol = 3) +
      labs(title = "Monthly Total Daily closing",
           subtitle = "Data plotted by year",
           y = "Closing trend",
           x = "Month") + theme_bw(base_size = 15) +
  scale_x_date(date_labels = "%b")
```
From this, The year has been highest performing in quarters 2 and 3.

Creating new dataframe with Date and Closing value
```{r}
PFizer %>% select(Date,Close)
```
Since we have made many of our analysis from the data, we will now be forecasting the data for the coming year(s).

To check if the dataset is considered as time series by R Studio
```{r}
class(PFizer)
```
It says it is data frame. It should be time series data frame so that it can be used to forecast using time series model.

```{r}
PFizer_time <-ts(PFizer$Close,start=min(PFizer$Date),end=max(PFizer$Date),frequency = 1)
```

```{r}
class(PFizer_time)
```

```{r}
library(forecast)
```


```{r}
plot(PFizer_time)
```
To check if our data is stationary or not and is eligible for forecasting, we will find the auto correlation function.
```{r}
acf(PFizer_time)
```
Here, we see that the values are shooting higher than the blue dotted horizontal lines which says that they are not stationary.

Also, let's check the partial auto correlation function.
```{r}
pacf(PFizer_time)
```
Here, the partial autocorrelation function doesn't show much out of range values.

Let's do Augmented Dickey-Fuller Test to test the stationarity one last time.
```{r}
adf.test(PFizer_time)
```
Form the above test, we see that the p-value is not less than or equal to 0.05 which says that the data is definitely not stationary.

To fit a good and best ARIMA model, we will use auto.arima function to run all possible sets and choose the best fitting model.
```{r}
#Fit an ARIMA model

fit_arima<-auto.arima(PFizer_time,ic="aic",trace = TRUE)

```
After running the trace, we see that ARIMA(1,1,0) is the best fitting model.

Let's check the residuals and graphs of ARIMA model.
```{r}
print(summary(fit_arima))
checkresiduals(fit_arima)
```
We will now be using our ARIMA model to forecast for next five years with confidence interval as 95%.

```{r}
#Forecast with ARIMA model
fcst<-forecast(fit_arima,level=c(95),h=5)
```

```{r}
plot(fcst,include = 10)
```
Testing the result of our forecast.
```{r}
Box.test(fcst$resid,lag=5,type="Ljung-Box")
```

We see here that the p-value is not below 0.05 and hence it says that our model doesn't have much correlation and is a good forecast.

